---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = file.path("man", "figures", "README-"),
  out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License](https://img.shields.io/badge/license-GPL--3-blue.svg?style=flat)](https://spdx.org/licenses/GPL-3.0-only.html)
[![Release](https://img.shields.io/github/release/inbo/flandersqmd.svg)](https://github.com/inbo/flandersqmd/releases)
![GitHub](https://img.shields.io/github/license/inbo/flandersqmd)
[![R build status](https://github.com/inbo/flandersqmd/workflows/check%20package%20on%20main/badge.svg)](https://github.com/inbo/flandersqmd/actions)
![r-universe name](https://inbo.r-universe.dev/badges/:name?color=c04384)
![r-universe package](https://inbo.r-universe.dev/badges/flandersqmd)
[![Codecov test coverage](https://codecov.io/gh/inbo/flandersqmd/branch/main/graph/badge.svg)](https://app.codecov.io/gh/inbo/flandersqmd?branch=main)
![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/inbo/flandersqmd.svg)
![GitHub repo size](https://img.shields.io/github/repo-size/inbo/flandersqmd.svg)
<!-- badges: end -->

# flandersqmd

The goal of `flandersqmd` is to provide a set of tools to create reports using the `flandersqmd` family of quarto extensions.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pkg_install("inbo/flandersqmd")
```

A stable version is available on [r-universe](https://inbo.r-universe.dev/):
``` r
install.packages("flandersqmd", repos = c("https://inbo.r-universe.dev", "https://cloud.r-project.org"))
```

## Example

### Working on a book style report

A book style report provides a structure to write a report in a book style format.
It is based on the `quarto` book format and uses the `flandersqmd-book` extension.
The available output formats are pdf and html.

Though optional, we strongly recommend to create a `checklist` project first.
For more information on how to create a `checklist` project, see the [`checklist`](https://inbo.github.io/checklist/articles/getting_started_project.html) package documentation.

```{r prepare-checklist, eval = FALSE}
library(checklist)
# Where to store the checklist project
path <- tempfile()
dir.create(path)
# create the checklist project
create_project(path = path, project = "flandersqmd-book")
```

Then generate a skeleton for the report using `create_report()`.
The function will guide you interactively through the process of setting up a report.
Once the setup is completed the report will automatically be rendered in both html and pdf format.
The respective output files are located in the output folder of the report folder.

```{r prepare-skeleton, eval = FALSE}
# Create a report skeleton
library(flandersqmd)
create_report(file.path(path, "flandersqmd-book"), shortname = "myreport")
```

When starting a new report, you often don't have all the metadata available yet.
Run `insert_missing_metadata()` to add the required metadata to the report when it becomes available to you.
E.g. the report number, DOI, ...

```{r insert-missing-metadata, eval = FALSE}
insert_missing_metadata(file.path(path, "flandersqmd-book", "myreport"))
```

`flandersqmd` does not create a cover page.
You have to provide a cover page yourself.
Use `add_cover()` to add a cover page to the report and then render the report again to include the cover page.
Note that you can render a draft version of the report without cover page or missing metadata.

```{r add-cover, eval = FALSE}
add_cover(
  report_path = file.path(path, "flandersqmd-book", "myreport"),
  cover_pdf = "path/to/cover.pdf"
)
# Render the report
quarto::quarto_render(file.path(path, "flandersqmd-book", "myreport"))
```
