# `flandersqmd`?

## What is Quarto?

- Quarto is the successor of bookdown
- Quarto is a more modern and flexible framework
- It allows to combine code and text in a single source
- It can generate documents in multiple formats from the same source
    - pdf, HTML, epub, ...
- Quarto supports R, Python, Julia and Observable JavaScript

## What is `flandersqmd`?

- `flandersqmd` is an R package that helps to set-up and render Quarto reports
- it provides templates and functions to create reports in the Flemish government style
- it is inspired by the `INBOmd` package for Rmarkdown/bookdown reports

## Why `flandersqmd` instead of `INBOqmd`?

- we could only reuse a part of the code from `INBOmd`
- the Flemish government has a common style guide for all entities
- `R` is becoming more common in other entities
- `flandersqmd` parametrised the entity specific settings
- `flandersqmd`
    - clear distinction with `INBOmd`
    - allows other entities to use the same framework

## Available extensions

- `flandersqmd-book`: report in pdf and or HTML format
- `flandersqmd-website`: HTML format for pages
    - recommended for reports with a lot of interactive figures
- `flandersqmd-revealjs`: slides in HTML format
- `flandersqmd-leaflet`: Brochure-like report without title page and colophon

# Structure of a Quarto project

## Files 

- content
    - text only files: `*.md`
    - files with text and code: `*.qmd`
- configuration file: `_quarto.yml`
- references: `*.bib`

## Input and output files

- require main `index.md` or `index.qmd` file is required.
- quarto requires you to list the files in the `_quarto.yml` file
    - list them in the relevant order in the `chapters` section under the `book` section
- for HTML output
    - Quarto converts each input file into one output file with the same name

## Formats

- Quarto install the extensions into the project
    - you need a copy of the extension in every project
    - the copy in a project is independent of the extension source
    - you need to update the extension manually in every project

## Render strategies

- Quarto uses the knit and merge strategy
    - pro: chapters can be rendered independently
    - contra: you need to store and load objects to reuse them

## Caching

- cache: stores the results of a code chunk
    - reuses the cached results if the **code** did not change
    - note that changes in data or packages are not detected
- freeze: stored the rendered output of a file
    - reuses the stored output if the **file** did not change
    - renders only the files that changed

## Rendering the report in RStudio GUI

- Create an RStudio project in the root folder of the report
- go to the `Build` pane
- click the `Render book` button
    - renders the selected format (HTML and/or pdf)
- you can also render a single chapter by clicking the `Render` button in the editor toolbar
    - only works for HTML output

## Rendering the report in the terminal

- `quarto render` to run all formats
- `quarto render --to flandersqmd-book-pdf` to render only the pdf format
- `quarto render --to flandersqmd-book-html` to render only the HTML format
- `quarto preview` to preview the HTML output in your browser

## Code chunk options

- Rmarkdown style (options in chunk header) still works
- Quarto style is preferred
    - Options in the chunk, starting with `#|`
    - Options can span multiple lines
        - Start extra lines with two extra spaces
    - Captions can contain markdown formatting and special characters

```r
#| echo: false
#| fig-cap: This is a _long_ figure caption that
#|   spans **multiple** lines
```

## Dynamic chunk options

- You can use `!expr` to set chunk options dynamically
- The expression is evaluated in `R` and the result is used as the option value

```r
#| eval: !expr 'nrow(observed) > 0'
#| fig.cap: !expr 'sprintf(toestand, params$name)'
```

## Cross-references

- Quarto uses `@chunk-label` to refer to a chunk
- the prefix determines the type of cross-reference
    - `fig-:` for figures
    - `tbl-:` for tables
    - `sec-:` for sections
    - `eq-:` for equations
- You need to set the figure and table captions with the `fig-cap` and `tbl-cap` options

# Getting started with `flandersqmd`

## The easy way: the `flandersqmd` package

- functions to set up a report interactively
- set a proper `_quarto.yml` file
- currently only supports the book format

## The easy way: the `flandersqmd` package

1. `flandersqmd::create_report()` generates a skeleton
1. Write the report content
1. Request the required metadata (report number, DOI, ...)
1. `flandersqmd::insert_missing_metadata()` adds missing metadata
1. Request a cover
1. `flandersqmd::add_cover()` adds the cover page

## The easy way: the `flandersqmd` package

- `flandersqmd::add_author()`: add an author or reviewer to the report
- `flandersqmd::add_address()`: add a client or cooperation partner to the report
- `flandersqmd::add_abstract()`: add an abstract to the report
- `flandersqmd::add_recommendations()`: add recommendations to the report
- `flandersqmd::add_chapter()`: add a chapter to the report
- `flandersqmd::add_bibliography()`: add a bibliography to the report
- `flandersqmd::inbo_website()`: create a zip file for the INBO website

## The hard way: manually

Full instructions are available in the [`flandersqmd-book` documentation](https://github.com/inbo/flandersqmd-book){target="_blank"}.

1. Create an empty Quarto book project
1. Install the extension with `quarto install extension inbo/flandersqmd-book`
1. Update the `_quarto.yml` as instructed in the [documentation](https://github.com/inbo/flandersqmd-book){target="_blank"}

The documentation provides source code for all available format options.

# `flandersqmd` section in `_quarto.yml`

## Selection of corporate identity style

- `entity`: the name of the entity.
- `level`: the level of the corporate identity style.
    - 1: Flanders
    - 2: entity
- `public_report`: whether the report is a public report or not.
- `colophon`: whether to display a colophon.
- `shortname`: a short name for the report, used in the file names

## Mandatory fields

:::: {.columns}

::: {.column width="40%"}
- `title`
- `author`
- `reviewer`
- `year`
- `reportnr`
:::

::: {.column width="40%"}
- `cover`
- `coverphoto`
- `coverdescription`
- `doi` ^[Only public report]
- `depotnr` <sup>1</sup>
:::

::::

## Optional fields

:::: {.columns}

::: {.column width="40%"}
- `client`
- `clientlogo`
- `clienturl`
- `subtitle`
- `ordernr`
:::

::: {.column width="40%"}
- `cooperation`
- `cooperationlogo`
- `cooperationurl`
- `watermark`
:::

::::

# Demo

## Create a report skeleton

1. Generate a skeleton with `create_report()`.
1. Add a `R` figure and table and a cross-reference to them.
1. Add missing metadata with `insert_missing_metadata()`.
1. Add the cover page with `add_cover()`.

# Parametrised reports

## Prepare the report

- Add parameters to the report to make it reusable for different analyses.
- Place it in the yaml front matter of every file.
- Access the parameters in the code chunks with `params$parameter_name`.

```
---
params:
  data: "data"
  location: 20502
  name: Fort Lier (3010)
---
```

## Render the report with different parameters

- Create a data frame with the different parameter sets.
- Loop over the rows of the data frame.

```{r}
#| echo: true
#| eval: FALSE
#| label: render-parameters
for (i in seq_len(nrow(to_do))) {
  quarto_render(
    ".", quiet = FALSE, as_job = FALSE,
    execute_params = list(name = to_do$name[i])
  )
}
```

# Child documents in quarto

## Child documents

- A child document is a single macro file that you can reuse multiple times in the document
    - e.g. a standard analysis repeated for several species, region, variables, ...
- Unlike bookdown, quarto does not support child documents out of the box
- We still can use bookdown machinery in a work around

## Macro `qmd` file

- Create a working `qmd` file for a single analysis
- Make sure you can do the selection based on the parameters
- Add one or more macro variables by using the `{{macro_name}}` syntax
    - `{{macro_name}}` will be replaced by the value of the variable
- Use `knitr::knit_expand()` to replace the macro variables with their values
- Write the result to a `qmd` file

## Workflow

1. Create a project with the macro `qmd` files
1. Create a render script that
    - reads the macro variables from a data frame
    - loops over the rows of the data frame
    - calls `knitr::knit_expand()` to replace the macro variables with their values
    - writes the result to a `qmd` file in a different quarto folder
    - update the `_quarto.yml` in the other quarto folder

## Example macro file

```
---
title: "Analysis for {{species}}"
----

{{vernacular}}
```

## Example render script

```{r}
#| echo: true
#| eval: false

todo <- data.frame(
  species = c("Erinaceus europaeus", "Talpa europaea"),
  vernacular = c("European hedgehog", "European mole"),
  file = c("hedgehog.qmd", "mole.qmd")
)
for (i in nrow(todo)) {
  knitr::knit_expand(
    "macro.qmd", species = todo$species[i],
    vernacular = todo$vernacular[i]
  ) |>
    writeLines(todo$file[i])
}
```
